---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    df_print: paged
    toc: true
    toc_float: false
    code_folding: hide
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load necessary packages
```{r loading, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(skimr)
library(forcats)
```



# load data
```{r}
ilo_working_hours <- get(load("Data/ilo_working_hours.RData"))
ilo_hourly_compensation <- get(load("Data/ilo_hourly_compensation.RData"))
```


# explore your dataset
```{r}
glimpse(ilo_working_hours)
glimpse(ilo_hourly_compensation)
```

```{r}
ilo_data <- ilo_working_hours %>% 
  inner_join(ilo_hourly_compensation, c("country","year"))
```

```{r}
# I will reverese the two datasets to see whether they differ or not
ilo_hourly_compensation %>% 
  inner_join(ilo_working_hours, c("country","year"))
```

Well. Both outputs are identical. It looks like that it doesn't matter.

```{r}
# Count the resulting rows
ilo_data %>% 
  count()
```

```{r}
#examine ilo_data
ilo_data
```

```{r}
#Turn year and country into a factor
ilo_data_corrected <- ilo_data %>% 
  mutate(year = as.factor(as.numeric(year)),
         country = as.factor(country))
# examine changes
glimpse(ilo_data_corrected)
```

```{r}
# Select European countries
european_countries <- c("France", "Italy","Norway","Finland","Spain", "Sweden","United Kingdom","Belgium","Ireland","Luxembourg","Portugal","Netherlands","Germany","Hungary","Austria", "Czech Rep.","Switzerland")

#check the length of European countries (it should be 17)
length(european_countries)
ilo_eu <- ilo_data_corrected %>% 
  filter(country %in% european_countries)
```

```{r}
# examine your data
str(ilo_eu)
```

```{r}
# do some calculations by year
ilo_eu %>% 
  group_by(year) %>% 
  summarize(mean_hourly_compensation = mean(hourly_compensation),
            mean_working_hours = mean(working_hours))
```

```{r}
# filter for 2006
plot_data <-  ilo_eu %>% 
  filter(year == 2006)

# Create the scatter plot and adding necessary lables
# then save it in order to make further edits and customizations
ilo_plot <- ggplot(plot_data) +
  geom_point(aes(x = working_hours, y = hourly_compensation)) +
  labs(title = "The more people work, the less compensation they seem to receive",
      x = "Working hours per week",
      y = "Hourly compensation",
      subtitle = "Working hours and hourly compensation in European countries, 2006",
      caption = "Data source: ILO, 2017"
  )

#check ilo_plot
ilo_plot
```

```{r}
# Try out the minimal theme
ilo_plot + theme_minimal()
```

```{r}
ilo_plot + theme_bw()
```

```{r}
# Let's make more customization to the text
ilo_plot <- ilo_plot + theme_minimal() + 
  theme(text = element_text(family = "serif"),
        plot.title = element_text(color = "gray25",face = "bold", size = 15),
        plot.caption = element_text(color = "blue", face = "bold"),
        plot.subtitle = element_text(size = 12))

# observe
ilo_plot
```

```{r}
# We can even do more customization to our plot (background and dimension)
ilo_plot + theme(
  plot.background= element_rect(fill = "gray95"),
  plot.margin = unit(c(5,10,5,10), units = "mm")
)
```

# Prepare the data set for the faceted plot

You're now going to prepare your data set for producing the faceted scatter plot in the next exercise, as mentioned in the video.
For this, the data set needs to contain only the years 1996 and 2006, because your plot will only have two facets. ilo_data has been pre-loaded for you.


```{r}
ilo_data_96_06 <-ilo_data_corrected %>%
  filter(year == 2006 | year == 1996)
```

```{r}
# Put your variables
ggplot(ilo_data_96_06, aes(x = working_hours, y = hourly_compensation, color = year)) + geom_point() + 
labs( #add your lables' details here
      title = "The more people work, the less compensation they seem to receive",
      subtitle = "Working hours and hourly compensation in European countries,2006",
      caption = "Data source: ILO, 2017",
      x = "Working hours per week",
      y = "Hourly compensation"
  ) + # use facet_grid()
facet_grid(. ~ year)

# let's save our plot to a variable
plot1 <- ggplot(ilo_data_96_06, aes(x = working_hours, y = hourly_compensation, color = year)) + geom_point() + 
labs( #add your lables' details here
      title = "The more people work, the less compensation they seem to receive",
      subtitle = "Working hours and hourly compensation in European countries,2006",
      caption = "Data source: ILO, 2017",
      x = "Working hours per week",
      y = "Hourly compensation"
  ) + # use facet_grid()
facet_grid(. ~ year)
```

```{r}
# Make your own theme
theme_ilo <- function(){
  theme_minimal() +
    theme(
    plot.background = element_rect(fill = "gray95"),
    text = element_text(family = "serif", colour = "gray25"),
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(color ="blue", face = "bold"),
    plot.margin = unit(c(5,10,5,10), units = "mm"),
    plot.subtitle = element_text(size = 12)
  )
}
```

```{r}
# test your theme
plot1 + theme_ilo()
```

```{r}
# let's make one more customization
plot1 + theme_ilo() + theme_bw() + theme(text = element_text(family = "serif"), plot.title = element_text(face = "bold"),
  strip.background = element_rect(color= "black", fill = "gray60"),
  strip.text = element_text(color ="white",face = "bold", family = "serif")
)
```

# How to create dot plots (interesting one!) using geom_path
```{r}
# create a dot plot using ilo_data_corrected
ggplot(ilo_data_96_06, aes(x = working_hours, y= country)) + geom_path() + theme_minimal()
```

```{r}
# let's add an arrow to each path (we must only use two points in time)
ggplot(ilo_data_96_06) + 
  geom_path(aes(x=working_hours, y = country), arrow = arrow(length = unit(1.5,"mm"), type = "closed")) + geom_text(
          aes(x = working_hours,
              y = country,
              label = round(working_hours, 1))
        )
```

# Next time I will polish the dot plot
```{r}
# To arrange countries according to working hours, we need to use `forcats` package and reorder the `country` factor
ilo_data_96_06 <- ilo_data_96_06 %>% 
  arrange(year) %>% 
  mutate(country = fct_reorder(country, working_hours, min))

# test your changes
ilo_data_96_06$country

# Plot again
ggplot(ilo_data_96_06) + 
  geom_path(aes(x=working_hours, y = country), arrow = arrow(length = unit(1.5,"mm"), type = "closed")) + geom_text(
          aes(x = working_hours,
              y = country,
              label = round(working_hours, 1))
        )

```

```{r}

# let's plot again after adjusting lables
plot_object <- ggplot(ilo_data_96_06) + 
  geom_path(aes(x=working_hours, y = country), arrow = arrow(length = unit(1.5,"mm"), type = "closed")) + geom_text(
          aes(x = working_hours,
              y = country,
              label = round(working_hours, 1),
              hjust = ifelse(year == "2006",1.4,-0.4)),
          family = "serif", colour = "blue",size = 2
        )
```

As you can see, there is a lot of overlapping in the graph. We will solve this using either `coord_cartesian` or `xlim` and `ylim` functions.

```{r}
#let's make more adjustments with dimensions
ilo_dot_plot <- plot_object + coord_cartesian(xlim = c(20,80)) + 
labs( #add your lables' details here
      title = "The more people work, the less compensation they seem to receive",
      subtitle = "Working hours and hourly compensation in European countries,2006",
      caption = "Data source: ILO, 2017",
      x = "Working hours per week",
      y = "Hourly compensation"
  ) + theme_ilo() 
```

```{r}
#optimize for mobile devices
# Compute temporary data set for optimal label placement
median_working_hours <- ilo_data_corrected %>%
  group_by(country) %>%
  summarize(median_working_hours_per_country = median(working_hours)) %>%
  ungroup()

# Have a look at the structure of this data set
str(median_working_hours)

ilo_dot_plot +
  # Add label for country
  geom_text(data = median_working_hours,
            aes(y = country,
                x = median_working_hours_per_country,
                label = country),
            vjust = 2,
            family = "serif",
            color = "gray25") +
  # Remove axes and grids
  theme(
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    # Also, let's reduce the font size of the subtitle
    plot.subtitle = element_text(size = 9)
  )
```


