---
title: "Categorical data with tidyverse" 
subtitle: "Provisional file"
author: "Insert your name here"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    df_print: paged
    toc: true
    toc_float: false
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load necessary packages
```{r loading, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(skimr)
library(forcats)
library(tidyr)
library(stringr)
library(scales)
```

# Explore your dataset

## load dataset and take a glimpse
```{r}
multiple_choice_responses <- read_csv("Data/smc_with_js.csv")

# have a glimpse
skim(multiple_choice_responses)
```

## Check for factors
```{r}
is.factor(multiple_choice_responses$CurrentJobTitleSelect)
```

## Convert all the character variables into factors
```{r}
responses_as_factors <- multiple_choice_responses %>% 
  mutate_if(is.character,as.factor)
```

## determine the number of levels of each factor
```{r}
number_of_levels <- responses_as_factors %>% 
  summarise_all(nlevels) %>% 
  gather(variable,num_levels)
```

## Select the 3 rows with the highest number of levels

```{r}
number_of_levels %>% 
  top_n(3,num_levels)
```

```{r}
## Filtering for the variable `CurrentJobTitleSelect`, pull the number of levels it has.

number_of_levels %>% 
  filter(variable == "CurrentJobTitleSelect") %>%
  pull()
```

## `pull()` the column `CurrentJobTitleSelect.` and get the values of the levels.
```{r}
responses_as_factors %>% 
  pull(CurrentJobTitleSelect) %>%
  levels()
```

## reordering bar charts by frequency
```{r}
responses_as_factors %>% 
  filter(!is.na(CurrentJobTitleSelect)) %>% 
ggplot(aes(x = fct_infreq(CurrentJobTitleSelect))) + geom_bar(fill = "orange", color = "black") + coord_flip() + theme_minimal()
```

## Try reversing the order 
```{r}
responses_as_factors %>% 
  filter(!is.na(CurrentJobTitleSelect)) %>% 
ggplot(aes(x = fct_rev(fct_infreq(CurrentJobTitleSelect)))) + geom_bar(fill = "orange", color = "black") + coord_flip() + theme_minimal()
```

## Changing the order of factor levels
```{r}
# Get the levels of WorkInternalVsExternalTools
levels(responses_as_factors$WorkInternalVsExternalTools)

# Reorder the levels from internal to external 
responses_as_factors <- responses_as_factors %>% 
  mutate(WorkInternalVsExternalTools = fct_relevel(WorkInternalVsExternalTools, "Entirely internal", "More internal than external", "Approximately half internal and half external", "More external than internal","Entirely external" , "Do not know"))
```

```{r}
# Check
levels(responses_as_factors$WorkInternalVsExternalTools)
```

## Make a bar plot of the responses
```{r}
responses_as_factors %>% 
  filter(!is.na(WorkInternalVsExternalTools)) %>% 
ggplot(aes(x = WorkInternalVsExternalTools)) + geom_bar() + coord_flip()
```

## Tricks of `fct_relevel()`
```{r}
# Pick another variable
levels(responses_as_factors$FormalEducation)
```

```{r}
# Move "I did not complete any formal education past high school" and "Some college/university study without earning a bachelor's degree" to the front. 

responses_as_factors <- responses_as_factors %>% 
  mutate(FormalEducation = fct_relevel(FormalEducation,
                                       "I did not complete any formal education past high school", "Some college/university study without earning a bachelor's degree")) %>% 
  # Move "I prefer not to answer" to be the last level.
  mutate(FormalEducation = fct_relevel(FormalEducation, "I prefer not to answer", after = Inf)) %>% 
  # Move "Doctoral degree" to be the sixth level (after the fifth level).
  mutate(FormalEducation = fct_relevel(FormalEducation, "Doctoral degree", after = 5))
```

```{r}
# Examine
levels(responses_as_factors$FormalEducation)
```

## Renaming a few levels
```{r}
# make a bar plot of the frequency of FormalEducation
ggplot(responses_as_factors, aes(x= FormalEducation)) + geom_bar()
```

```{r}
# Now, rename "I did not complete any formal education past high school" to "High school" and "Some college/university study without earning a bachelor's degree" to "Some college".

responses_as_factors <- responses_as_factors %>% 
  mutate(FormalEducation = fct_recode(FormalEducation,
                                      "High school" = "I did not complete any formal education past high school",
                                      "Some college" = "Some college/university study without earning a bachelor's degree"))

# Test again and create a new bar plot
ggplot(responses_as_factors, aes(x= FormalEducation)) + geom_bar() + coord_flip()

```

## Collapsing factor levels
```{r}
# Examine the factor CurrentJobTitleSelect
levels(responses_as_factors$CurrentJobTitleSelect)

```

Collapse the levels of `CurrentJobTitleSelect` into a new variable, `grouped_titles`.
```{r}
responses_as_factors %>% 
  mutate(grouped_titles = fct_collapse(CurrentJobTitleSelect,
                                       "Computer Scientist" = c("Programmer", "Software Developer/Software Engineer"), 
        "Researcher" = "Scientist/Researcher", 
        "Data Analyst/Scientist/Engineer" = c("DBA/Database Engineer", "Data Scientist", 
                                              "Business Analyst", "Data Analyst", 
                                              "Data Miner", "Predictive Modeler"))) %>% 
  # Then take grouped_titles and put everything that isn't one of those three grouped titles into "Other Title". 
  mutate(grouped_titles = fct_other(grouped_titles, keep = c("Computer Scientist", 
                                     "Researcher", 
                                     "Data Analyst/Scientist/Engineer"))) %>% count(grouped_titles)
```

## lumping variables
```{r}
responses_as_factors %>%
  # remove NAs of MLMethodNextYearSelect
  filter(!is.na(MLMethodNextYearSelect)) %>%
  # create ml_method, which lumps all those with less than 5% of people into "Other"
  mutate(ml_method = fct_lump(MLMethodNextYearSelect, prop = 0.05)) %>%
  # count the frequency of your new variable, sorted in descending order
  count(ml_method, sort = TRUE)
```

## Preserving the most common levels
```{r}
responses_as_factors %>%
  # remove NAs 
  filter(!is.na(MLMethodNextYearSelect)) %>%
  # create ml_method, retaining the 5 most common methods and renaming others "other method" 
  mutate(ml_method = fct_lump(MLMethodNextYearSelect, n = 5, other_level = "other method")) %>%
  # count the frequency of your new variable, sorted in descending order
  count(ml_method, sort = TRUE)
```

## Grouping and reshaping similar columns
```{r}
# Select only the columns with `LearningPlatformUsefulness` in the name.
learning_platform_usefulness <- responses_as_factors %>% 
  select(contains("LearningPlatformUsefulness")) %>% 
  # Change the data from wide to long format with two columns, learning_platform and usefulness.
  gather(learning_platform, usefulness) %>% 
  # Remove rows where usefulness is NA.
  filter(!is.na(usefulness)) %>% 
  # Remove "LearningPlatformUsefulness" from each string in learning_platform.
  mutate(learning_platform = str_remove(learning_platform, "LearningPlatformUsefulness"))
```

```{r}
# examine your data
glimpse(learning_platform_usefulness)
```

## Summarizing data
```{r}
# Use count() to change the dataset to have one row per learning_platform usefulness pair with a column that is the number of entries with that pairing.

learning_platform_usefulness %>% 
  count(learning_platform, usefulness)
```

```{r}
   # Use add_count() to create a column with the total number of answers to that learning_platform.
learning_platform_usefulness %>% 
  count(learning_platform, usefulness) %>% 
  add_count(learning_platform, wt = n, name="nn")

```


```{r}
#Create a new column, perc, that's the percent of people giving a certain answer for that question. Save everything as a new dataset, perc_useful_platform.

(perc_useful_platform <- learning_platform_usefulness %>% 
  count(learning_platform, usefulness) %>% 
  add_count(learning_platform, wt = n, name="nn") %>% 
  mutate(perc = n/nn))
```

```{r}
# For each learning platform, create a line graph with usefulness on the x-axis and percentage of responses within the learning platforms on the y-axis.

ggplot(perc_useful_platform, aes(x = usefulness, y = perc, group = learning_platform, fill = usefulness)) + geom_col() + facet_wrap(~learning_platform) + theme_bw() + theme(
  axis.ticks.x = element_blank(), axis.text.x = element_blank()
)
```

## Creating an initial plot
```{r}
# Using the dataset learning_platform_usefulness, change usefulness to equal 0 if someone answered "Not Useful" and 1 otherwise.
learning_platform_usefulness %>% 
  mutate(usefulness = ifelse(usefulness == "Not Useful",0,1))
```

```{r}
usefulness_by_platform <- learning_platform_usefulness %>% 
 mutate(usefulness = ifelse(usefulness == "Not Useful",0,1)) %>%
  # Group the data by each platform.
  group_by(learning_platform) %>% 
# Create a new column, avg_usefulness, the mean usefulness of each platform.
  summarise(avg_usefulness = mean(usefulness))
# Save the result as a new dataset, usefulness_by_platform

#examine the dataset
usefulness_by_platform
```


## Tricks of ggplot2

### Editing plot text
```{r}
# Create a scatter plot of avg_usefulness by learning_platform, using the dataset you created, usefulness_by_platform.
ggplot(usefulness_by_platform, aes(y = avg_usefulness, x = learning_platform)) + geom_point() + theme(
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, face = "bold")
) + labs(x = "Learning Platform", y = "Percent finding at least somewhat useful") +
  # Change the y-axis scale to be a percentage
  scale_y_continuous(labels = scales::percent)
```
2363137
2363314
2362483
### Reordering graphs (Ascendingly)
```{r}
# Order learning_platform in the graph by avg_usefulness so that, from left to right, it goes from highest usefulness to lowest.

ggplot(usefulness_by_platform, aes(y = avg_usefulness, x = fct_reorder(learning_platform, avg_usefulness))) + geom_point() + theme(
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, face = "bold")
) + labs(x = "Learning Platform", y = "Percent finding at least somewhat useful") +
  # Change the y-axis scale to be a percentage
  scale_y_continuous(labels = scales::percent)
```

### Ordering graphs (descendingly)
```{r}
# Order learning_platform in the graph by avg_usefulness so that, from left to right, it goes from highest usefulness to lowest.

ggplot(usefulness_by_platform, aes(y = avg_usefulness, x = fct_rev(fct_reorder(learning_platform, avg_usefulness)))) + geom_point() + theme(
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, face = "bold")
) + labs(x = "Learning Platform", y = "Percent finding at least somewhat useful") +
  # Change the y-axis scale to be a percentage
  scale_y_continuous(labels = scales::percent)
```

## Changing and creating variables with case_when()
### case_when() with a single variable
```{r}
# filter for rows where Age is between 10 and 90
responses_as_factors %>% 
  filter(Age >= 10 & Age <= 90) %>% 
  # Create the generation variable based on age 
  mutate(generation = case_when(
    between(Age,10,22) ~ "Gen Z",
    between(Age,23,37) ~ "Gen Y",
    between(Age,38,52) ~ "Gen X",
    between(Age,53,71) ~ "Baby Boomer",
    between(Age,72,90) ~ "Silent",
  )) %>% 
  count(generation)
```

### case_when() from multiple variables
```{r}
responses_as_factors %>% 
# Filter out people who selected Data Scientist as their Job Title
  filter(!(CurrentJobTitleSelect == "Data Scientist"))
```

```{r}
# Create a new variable, job_identity, based on their current job title and whether they fully identify as a data scientist.
responses_as_factors %>%
  # Filter out people who selected Data Scientist as their Job Title
    filter(CurrentJobTitleSelect != "Data Scientist") %>%
    # Create a new variable, job_identity
    mutate(job_identity = case_when(
        CurrentJobTitleSelect == "Data Analyst" & 
        DataScienceIdentitySelect == "Yes" ~ "DS analysts", 
        CurrentJobTitleSelect == "Data Analyst" & 
        DataScienceIdentitySelect %in% c("No", "Sort of (Explain more)") ~ "NDS analyst", 
        CurrentJobTitleSelect != "Data Analyst" & 
        DataScienceIdentitySelect == "Yes" ~ "DS non-analysts", 
        TRUE ~ "NDS non analysts")) %>% 
  mutate(JobSatisfaction=parse_number(as.character(JobSatisfaction))) %>% 
  group_by(job_identity) %>% 
  summarise(avg_js = mean(JobSatisfaction, na.rm = TRUE))

```

## Case study on Flight Etiquette

- loading data
```{r}
flying_etiquette <- read_csv("Data/flying-etiquette.csv")
```

- take a glimpse so you have all your variables in hands
```{r}
skim(flying_etiquette)
```


- Change all character columns into factor columns and remove people who responded "Never" to a question asking if they have flown before. 
```{r}
# useful trick here is mutate_if
flying_etiquette %>% 
  #change character variable to factor
  mutate_if(is.character, as.factor) %>% 
  # Filter out those who have never flown on a plane
  filter(`How often do you travel by plane?` != "Never")
```

- Select columns where "rude" is in the column name. 
- Change the dataset from "wide" to "long", with the variable names in a column called "response_var" and the values in a column called "value."
```{r}
# Copy the previous chunk and save your work to gathered_data
gathered_data <- flying_etiquette %>% 
  #change character variable to factor
  mutate_if(is.character, as.factor) %>% 
  # Filter out those who have never flown on a plane
  filter(`How often do you travel by plane?` != "Never") %>%
  # Select columns where "rude" is in the column name.
  select(contains("rude")) %>%
  # Change format from wide to long
  gather("response_var", "value")
```

- Check your data
```{r}
glimpse(gathered_data)
```

## Data preparation and regex
- Use `str_remove` to remove everything before and including "rude to " (with the space at the end) in the `response_var` column.

```{r}
gathered_data %>% 
  # Remove everything before and including "rude to " (with that space at the end!)
  mutate(response_var = str_remove(response_var,".*rude to ")) %>% 
  # Remove "on a plane"
  mutate(response_var = str_remove(response_var, " on a plane"))
```

- Remove rows with NA in the value column
- Create a new variable, rude, which is 0 if the value column is "No, not rude at all" or "No, not at all rude" and 1 otherwise.
```{r}
dichotimized_data <- gathered_data %>% 
  # Remove everything before and including "rude to " (with that space at the end!)
  mutate(response_var = str_replace(response_var,".*rude to ","")) %>% 
  # Remove "on a plane"
  mutate(response_var = str_replace(response_var, " on a plane","")) %>%
  # Remove rows with NA in the value column
  filter(!is.na(value)) %>%
  # Dichotomize the value variable to make a new variable, rude
  mutate(rude = if_else(value %in% c("No, not rude at all", "No, not at all rude"), 0,1))
  
```

- Check your data
```{r}
head(dichotimized_data,10)
```


- Summarise the data set into two columns, the question (i.e. `response_var`), and a new column, `perc_rude`, the mean of the `rude` column for each question.
- Save it as `rude_behaviors` and then view your new dataset.

```{r}
rude_behaviors <- gathered_data %>%
    mutate(response_var = str_replace(response_var, '.*rude to ', '')) %>%
    mutate(response_var = str_replace(response_var, 'on a plane', '')) %>%
	# Remove rows that are NA in the value column
	filter(!is.na(value)) %>%
    mutate(rude = if_else(value %in% c("No, not rude at all", "No, not at all rude"), 0, 1)) %>%
  # Group by response_var
  group_by(response_var) %>% 
  # Create perc_rude, the percent considering each behavior rude
  summarise(perc_rude = mean(rude))
```

- Examine your data
```{r}
rude_behaviors
```

## Receating the plot

- Order `response_var` by `perc_rude`
- Make a bar plot of `response_var` by `perc_rude`. Save it as `initial_plot`.

```{r}
initial_plot <- rude_behaviors %>% 
  ggplot(aes(x = fct_reorder(response_var, perc_rude), y = perc_rude)) + geom_col()

#View your plot
initial_plot
```

- To match the FiveThirtyEight graph:
  - Add its title, subtitle, and caption
  - Set the labels for x and y-axis as empty strings
  - Save it as a new plot, `titled_plot`.

```{r}
titled_plot <- initial_plot + 
  labs(title = "Hell Is Other People In A Pressurized Metal Tube",
       subtitle = "Percentage of 874 air-passenger respondents who said action is very or somewhat rude",
       caption = "Source: SurveyMonkey Audience", x="", y="")

titled_plot
```

- Flipping things around
  - Flip the x- and y-axis.
  - Remove the x-axis ticks.
```{r}
flipped_plot <- initial_plot + 
  coord_flip() + 
  # Remove the x-axis ticks and labels
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

flipped_plot
```

- Finalizing the chart
  - Apply `percent()` to `perc_rude` to add a label to each bar with the percent rude value, so it will display as "85%".
```{r fig.height=10, fig.width=20}
# adding labels to bar plots
flipped_plot + theme_bw() + geom_text(aes(label = percent(perc_rude),
                             y = perc_rude + 0.03),
                         position = position_dodge(0.9),
              vjust = 1, color = "blue", fontface = "bold") + labs(y = "% rudeness", x = "Participants' responses") + theme(text = element_text(family = "serif", colour = "gray25", size = 20, face = "bold"))
```
  
  