

```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(moderndive)
library(readr)
library(skimr)
library(janitor)
```


```{r}
bakeoff <- read_csv("Data/bakeoff.csv", skip = 1)
```

```{r}
# View bakeoff
bakeoff
```

```{r}
# Filter rows where showstopper is UNKNOWN
bakeoff %>% 
  lapply(showstopper, is.na)
```

```{r}
skim(bakeoff)
```

# How to parse date
```{r}
parse_date("17 August 2010", format = "%d %B %Y")
```

```{r}
desserts <- read_csv("Data/Desserts/desserts.csv", col_types = cols(uk_airdate = col_date(format = "%d %B %Y")))
```

```{r}
glimpse(desserts)
```


Sort desserts in descending order by uk_airdate. when did the last episode air in the UK?
```{r}
desserts %>% 
  arrange(desc(uk_airdate))
```

Add another line to cols() to cast the technical column as a number. You will get a parsing error! Read the red text carefully.

```{r}
desserts <- read_csv("Data/Desserts/desserts.csv", 
                     col_types = cols(technical = col_number(),
                     uk_airdate = col_date(format = "%d %B %Y")))
```

```{r}
# examine your problems
problems(desserts)
```

```{r reference chunk}
# The problem can be solved by including "N/A" within the na values in read_csv function
desserts <- read_csv("Data/Desserts/desserts.csv", 
                     col_types = cols(technical = col_number(),
                     uk_airdate = col_date(format = "%d %B %Y")),
                     na = c("","NA","N/A"))
```

```{r}
# check
glimpse(desserts)
```


```{r}
# Now, we will do some edits to `results` variable. We want to make it factor instead of character

#copy a previous code from chunk 12
desserts <- read_csv("Data/Desserts/desserts.csv", 
                     col_types = cols(technical = col_number(),
                     uk_airdate = col_date(format = "%d %B %Y"),
                     #add another argument inside cols function
                     result = col_factor(levels = NULL)),
                     na = c("","NA","N/A"))

# check if you changes were made correctly
glimpse(desserts)
```

```{r}
# rename the variable signature_nut 
desserts <- desserts %>% 
  mutate(nut = signature_nut) %>% 
  select(-signature_nut)
```

Count the distinct values in the `nut` variable, sorting the result based on the number of rows. How many bakes included filberts as the nut?

```{r}
desserts %>% 
  count(nut,sort = TRUE) %>% 
  mutate(perc = n/sum(n)*100)
```

What is a "filbert", you may ask? Recode "filbert" to "hazelnut", store the result in the same nut column, and count again.

```{r}
desserts %>% 
  mutate(nut = recode(nut, "filbert" = "hazelnut", "no nut" = NA_character_)) %>% 
  count(nut, sort = TRUE)
```

Create a new dummy variable called tech_win: it should be equal to 1 if the baker was ranked first in the technical challenge, and 0 if not. Remember that when recoding numeric variables, you need to put the old value in backticks, for example, `5` = 6.

```{r}
desserts <- desserts %>% 
  mutate(tech_win = recode(technical, `1` = 1, .default = 0))
```

Use count to compare the number of rows where technical is equal to 1 versus all distinct values in the new tech_win variable.
```{r}
desserts %>% 
  count(technical == 1, tech_win)
```

```{r}
messy_ratings <- read_csv("Data/Ratings/02.03_messy_ratings.csv")
```

```{r}
problems(messy_ratings)
```


    For series with 10 episodes, which showed the most growth in viewers from the premiere to the finale? Which showed the least?

```{r}
ratings %>% 
  filter(episodes == 10) %>% 
  mutate(growth = e10_viewers - e1_viewers) %>% 
  group_by(series) %>% 
  select(series, growth)
  
```

```{r}
ratings <- ratings %>% 
  mutate(bbc = recode(channel, "Channel 4"=0,.default = 1))
```

```{r}
ratings %>% 
  select(channel, everything())
```

```{r}
ratings %>% 
  select(ends_with("day"))
```

```{r}
glimpse(messy_ratings)
```

```{r}
viewers_7day <- ratings %>% 
  select(series, ends_with("_7day") )
```

```{r}
viewers_7day %>% 
  select(viewers_7day_ = ends_with("_7day"))
```

```{r}
ratings %>% 
  select(ends_with("_7day"),-ends_with("_28day"))
```

```{r}
ggplot(rating, aes(x = series, y= e1)) + geom_col()
```


```{r}
ratings %>% 
  gather(key = "episode", value = "viewers_7day", e1:e10, factor_key = TRUE ,na.rm = TRUE)
```

```{r}
ggplot(tidy_ratings, aes(x = episode, y = viewers_7day, fill = series)) + geom_col()
```

```{r}
ratings2 %>% 
  gather(key = "episode", value = "viewers", e1_7day:e10_28day, na.rm = TRUE) %>% 
  separate(episode, into = c("episode", "day")) %>% 
  mutate(episode = parse_number(episode), day = parse_number(day))
```

```{r}
ratings %>% 
  gather(key = "episode", value = "viewers", e1:e10, na.rm = TRUE) %>% 
  mutate(episode = parse_number(episode)) %>% 
  ggplot(aes(x = episode, y = viewers, group = series, color = series)) + geom_point() + facet_wrap(~series) + theme_minimal()
  
```

```{r}
bakers %>% 
  mutate(skill = case_when(
    star_baker > technical_winner ~ "super_start",
    star_baker < technical_winner ~ "high_tech",
    TRUE ~ "well_rounded"
  ))
```

```{r}
bakers_skill %>% 
  filter(star_baker == 0 & technical_winner == 0) %>% 
  count()
```

```{r}
x <- 4
class(x)
```

```{r}
y <- c(4,TRUE)
class(y)
```

```{r}
dim(y)
```

```{r}
 x <- list(2, "a", "b", TRUE)
x[[2]]
length(x[[2]])
```

```{r}
x <- c(3, 5, 1, 10, 12, 6)
x[x %in% 1:5] <- 0
x
```

